---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(1234)
```

# anon

<!-- badges: start -->
<!-- badges: end -->

The anon package provides comprehensive tools for anonymizing sensitive information in R objects, including data frames, variable labels, lists, and character/factor/numeric vectors.

## Installation

You can install the development version of anon from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("nash-delcamp-slp/anon")
```

## Loading the Package

```{r}
library(anon)
```

Example data:

```{r}
library(dplyr)
data("starwars", package = "dplyr")

a_new_hope_intro <- c(
  "It is a period of civil war.", 
  "Rebel spaceships, striking from a hidden base, have won their first victory against the evil Galactic Empire.",
  "During the battle, Rebel spies managed to steal secret plans to the Empire's ultimate weapon, the DEATH STAR, an armored space station with enough power to destroy an entire planet.",
  "Pursued by the Empire's sinister agents, Princess Leia races home aboard her starship, custodian of the stolen plans that can save her people and restore freedom to the galaxy...."
)
```

## Core Anonymization with `anon()`

The main `anon()` function and helper `anon_*()` support anonymization of various R objects.

### Vector Anonymization

```{r}
# Character sequence anonymization - preserves unique mappings
names_vector <- c("John", "Paul", "John", "Keith")
anon_id_chr_sequence(names_vector)

# Numeric sequence anonymization
anon_id_num_sequence(starwars$name) |> 
  head()

# Preserve distribution while anonymizing values
anon_num_preserve_distribution(starwars$height) |> 
  head()

# Group numeric values into ranges
anon_num_range(starwars$mass, n_breaks = 10) |> 
  head()
```

### Data Frame Anonymization

```{r}
anon_starwars <- starwars |> 
  anon(
    # Pattern-based text replacement
    pattern_list = list(
      DARK = c("empire", "imperials?", "sith"),
      LIGHT = c("jedi", "rebels?")
    ),
    # Column-specific anonymization rules
    df_variable_names = list(
      name = anon_id_chr_sequence,
      homeworld = ~ anon_id_chr_sequence(.x, start = "Planet ")
    ),
    # Class-based anonymization rules
    df_classes = list(
      integer = ~ anon_num_range(.x, n_breaks = 10),
      numeric = anon_num_preserve_distribution
    )
  )

glimpse(anon_starwars)
```

## Shiny App

`anon` includes a local Shiny app for reviewing the current R environment,
selecting objects for a report, cleaning text, and building anonymized prompt
context.

``` r
library(anon)
library(dplyr)

starwars <- dplyr::starwars
run_anon_app()
```

By default, the app loads the current environment into the inventory and starts
with all available objects selected for `Objects for report`. The `Refresh
environment` button is available if the session changes after launch.

## Global Options System

Many functions in the anon package reference global `anon.*` options that make it easy to configure default behaviors across workflows.

Use `anon_options()` as the central helper for these settings. It is a thin wrapper around base `options()` for the supported anon package options:

- `anon_options()` with no arguments returns the current supported `anon.*` values.
- `anon_options(...)` sets those same underlying global options and invisibly returns the previous values, matching the values returned by `options()`.
- `options(old)` can still be used to restore previously captured values.

You can still call `options()` directly, but the examples below use `anon_options()` so all supported package options are configured in one consistent way.

```{r}
anon_options()
```

### Setting Default Patterns & Replacements

Set `pattern_list` to define patterns that should always be replaced.

```{r}
anon_options(
  pattern_list = list(
    DARK = c("empire", "imperials?", "sith"),
    LIGHT = c("jedi", "rebels?")
  )
)
anon(a_new_hope_intro)
```

The default options are applied after provided arguments.

```{r}
anon(
  a_new_hope_intro,
  pattern_list = list("good guy" = "rebel")
)
```

Reset the option:

```{r}
anon_options(pattern_list = NULL)
```

### Setting Default Data Frame Behavior

Set `df_variable_names` and `df_classes` to automatically replace content with a constant or the result of a function.

```{r}
old_options <- anon_options(
  df_variable_names = list(
    name = anon_id_chr_sequence,
    homeworld = ~ anon_id_chr_sequence(.x, start = "Planet ")
  ),
  df_classes = list(
    integer = ~ anon_num_range(.x, n_breaks = 10),
    numeric = anon_num_preserve_distribution
  )
)

starwars |>
  anon() |>
  glimpse()
```

Reset the options:

```{r}
old_options <- anon_options(
  df_variable_names = NULL,
  df_classes = NULL
)
```

### Setting a Default Replacement

Set `default_replacement` to define a fallback replacement.

```{r}
old_options <- anon_options(default_replacement = "[HIDDEN]")

anon(
  a_new_hope_intro,
  pattern_list = list("empire", "imperials?", "sith", "jedi", "rebels?")
)
```

Reset the option:

```{r}
old_options <- anon_options(default_replacement = NULL)
```

### Setting Default NLP Entity Replacements

Set `nlp_default_replacements` to define replacement values for different entity types.
Use `nlp_default_replacements()` to create the list input.

```{r}
# View the default replacement values for all entity types
str(nlp_default_replacements())

# Customize specific replacements
my_replacements <- nlp_default_replacements(
  person = "[NAME]",
  org = "[GROUP]",
  gpe = "[LOCATION]",
  propn = "[PROPER_NOUN]"
)

old_options <- anon_options(nlp_default_replacements = my_replacements)

anon_nlp_entities(a_new_hope_intro)
```

Reset the option:

```{r}
old_options <- anon_options(nlp_default_replacements = NULL)
```

### Enabling Automatic NLP Entity Anonymization

Set `nlp_auto` to control which entity types are automatically anonymized.
Use `nlp_auto()` to create the list input.

```{r}
# View default auto-anonymization settings (all FALSE by default)
str(nlp_auto())

# Enable all entity types
str(nlp_auto(default = TRUE))

# Enable only specific entity types
my_auto_settings <- nlp_auto(person = TRUE, org = TRUE, gpe = TRUE, propn = TRUE)
old_options <- anon_options(nlp_auto = my_auto_settings)

anon(a_new_hope_intro)
```

Reset the option:

```{r}
old_options <- anon_options(nlp_auto = NULL)
```

### Setting Default Data Summary Examples

Set `example_values_n` and `example_rows` to control the default example content used by `anon_data_summary()`, `anon_report()`, and the Shiny app's initial data-summary settings.
Use `anon_example_rows()` to create the row-example specification.

```{r}
old_options <- anon_options(
  example_values_n = 3,
  example_rows = anon_example_rows(
    n = 2,
    key = "homeworld",
    method = "random",
    n_key_values = 2,
    seed = 11
  )
)

anon_data_summary(list(starwars = starwars))
```

Reset the options:

```{r}
old_options <- anon_options(
  example_values_n = NULL,
  example_rows = NULL
)
```

## Automatic NLP-Powered Anonymization

The anon package includes powerful NLP functions that can automatically detect and anonymize different types of named entities in text.

### Individual NLP Functions

The package provides specific functions for different entity types:

- `anon_nlp_entities()`: Detect and anonymize select entities. By default, all entities.
- `anon_nlp_people()`: Detect and anonymize person names
- `anon_nlp_organizations()`: Detect and anonymize organization names
- `anon_nlp_places()`: Detect and anonymize locations (cities, countries, etc.)
- `anon_nlp_dates()`: Detect and anonymize dates
- `anon_nlp_numbers()`: Detect and anonymize monetary amounts and percentages

```{r}
anon_nlp_entities(a_new_hope_intro)
anon_nlp_people(a_new_hope_intro)
anon_nlp_organizations(a_new_hope_intro)
anon_nlp_places(a_new_hope_intro)
anon_nlp_dates(a_new_hope_intro)
anon_nlp_numbers(a_new_hope_intro)
```

## Data Environment Summary

The `anon_data_summary()` function provides an anonymized overview of the data in your environment:

```{r}
anon_data_summary()
```

## Pattern Detection Warnings

When using pattern-based anonymization, anon will detect potential matches and report warnings unless `check_approximate = FALSE`.

```{r}
starwars |> 
  anon(list(
    "blonde", "bleu", "imperials"
  )) |> 
  glimpse()
```
